Leadership across cultures and contexts

### 2019-10-28 leadership_across_cultures_contexts.Rmd
Working on MS text.
Exploring Bayesian models.

Across human societies, the influence of leaders stands to substantially shape the nature of group structure and group outcomes. Simultaneously, social norms, pressures from competing leaders or rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda. The interactions between the strategies of individual leaders, pressures of followership, and cultural forces and group dynamics likely play an out-sized role in contributing to organizational variation across human populations including social and economic trajectories of populations [@wiessner_risk_1982], and cultural evolutionary processes [@henrich_big_2015]. Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies. Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced to a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., CITE], accompanied by a growing body of empirical results [e.g., CITE]. This approach as been informative. Drawing on the study of leadership across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015]. Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains. Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated.

Leaders are individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be. Leadership roles are also often associated with specific rights and responsibilities. Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the challenges specific groups face. Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor certain types of leaders, however there can be substantial variation in the qualities of leaders within these groups as well as and between them.

@@

Leadership as a phenomena is a universal feature of human groups; among all societies there are particular contexts in which one individual maintains greater influence. Also, adopting the broad definition of leadership given above, parents can be defined as leaders of nuclear families. Strong community leadership, however, is common cross-culturally.

@@

The geographic distribution of the cultures in our sample is displayed in Fig. 1, with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1). The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X. The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.

@@

```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8, fig.cap="Support for each coded variable. Blue circles: percent of text records with evidence of support. Red triangles: percent of cultures with evidence of support."}
plot.variable.support_costs_benefits
```

@@

```{r qualitiesPCplot, fig.width=9, fig.height=6, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first of three components."}
qualities_component1_plot
```

@@

```{r functionsPCplot, fig.width=9, fig.height=6, fig.cap="Logistic PCA loadings of coded leader function variables on the first of three components."}
functions_component1_plot
```

@@

To explore potential population variation in the *Prestige vs. Dominance* leader qualities component and the *Mediation vs. Organization* leader functions component, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture.

We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “huntergatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed. We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, levels of male aggression, and social stratification. Fifty of these cultures are the same as, or very similar to, cultures in the PSF.

>>>>

Across human societies, the influence of leaders stands to substantially shape the nature of group structure and group outcomes. Simultaneously, social norms, pressures from competing leaders or rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda. The interactions between the strategies of individual leaders, pressures of followership, and cultural forces and group dynamics likely play an out-sized role in contributing to organizational variation across human populations including social and economic trajectories of populations [@wiessner_risk_1982], and cultural evolutionary processes [@henrich_big_2015]. Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies. Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010 and @kaplan_evolutionary_2009], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016]. This approach as been informative. Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019]. Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains. Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated. Current outstanding questions for evolutionary leaderhip scholars include the degree to which the nature of leadership is shaped by population-level variation and the costs and benefits leaders and followers encur across distinct leadership systems.

Evolutionay scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be. Leadership roles are also often associated with specific rights and responsibilities. Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face. Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor leaders with specific qualities, however there can be substantial variation in the traits of leaders within these groups as well as between them.

@@

Leadership as a phenomena is a universal feature of human groups; among all societies there are particular contexts in which one individual maintains greater influence. Also, adopting the broad definition of leadership given above, parents can be defined as leaders of nuclear families. Strong community leadership, however, is common cross-culturally [@garfield_review_2019; @lewis_leaders_1974].

@@

The geographic distribution of the cultures in our sample is displayed in Figure X, with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1). The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X. The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.

@@

```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable. Blue circles: percent of text records with evidence of support. Red triangles: percent of cultures with evidence of support."}
plot.variable.support_costs_benefits
```

@@

```{r qualitiesPCplot, fig.width=9, fig.height=7, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first of three components."}
qualities_component1_plot
```

@@

```{r functionsPCplot, fig.width=9, fig.height=7, fig.cap="Logistic PCA loadings of coded leader function variables on the first of three components."}
functions_component1_plot
```

@@

To explore potential population variation in the *Prestige vs. Dominance* leader qualities component and the *Mediation vs. Organization* leader functions component, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture with these componets as outcome measures.

For predictor variables we incorporate additional population-level measures. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “huntergatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, dependence on agriculture, and social stratification. Fifty of these cultures are the same as, or very similar to, cultures in the PSF. Specifically, we use the SCCS measures of community size, population density, and cultural complexity. Using these measures as covariates allows us to determine the degree to which, an aggregate measure of cultural complexity, mode of subsistence, or measures of community size and population dispersal most strongly impact the qualities and functions of leaders. We also include our researcher coded measure of group structure (See Table X) as a predictor.

```{r, fig.height=20, fig.width=9}
qc_bayes_model_post_plot
```
